### 2013-05-04 doc/dissertation/document/backmatter/pn_equations.tex
adding pn fission source derivation

>>>>

\section{Eigenvalue Form of the $P_N$ Equations}
\label{sec:pn_eigenvalue}
In addition to fixed source problems, the $P_N$ equations may be
modified to solve eigenvalue problems by applying the same
discretization techniques to the fission source in
Eq~(\ref{eq:eigenvalue_transport}). Considering the multigroup
case, we first replace the integration over energy with a summation
over energy groups:
\begin{multline}
\frac{1}{k} \chi(E) \iint \nu \sigma_f(x,E') \psi(x,\hat{\Omega}',E')
d\Omega' dE' =\\ \frac{1}{k} \sum_0^G \chi^g \int \nu \sigma^{g'}_f(x)
\psi^{g'}(x,\hat{\Omega}') d\Omega'\:,
  \label{eq:mg_fission_source}
\end{multline}
where $G = N_g-1$ and $N_g$ is the number of energy groups. Next, we
expand the angular flux in the fission term by applying the expansion
in Eq~(\ref{eq:flux_expansion}) and remove the dependence on the
azimuthal angle:
\begin{multline}
\frac{1}{k} \sum_0^G \chi^g \int \nu \sigma^{g'}_f(x)
\psi^{g'}(x,\hat{\Omega}') d\Omega' =\\ \frac{1}{k} \sum_0^G \chi^g
\int_{-1}^{1} \nu \sigma^{g'}_f(x) \sum_{n=0}^\infty
(2n+1)P_n(\mu)\phi^{g'}_n(x) d\mu'\:.
  \label{eq:fission_flux_expansion}
\end{multline}
We can again use the the orthogonality property of the Legendre
polynomials by multplying by $P_0(\mu) = 1$:
\begin{equation}
  \frac{1}{k} \sum_0^G \chi^g
  \int_{-1}^{1} \nu \sigma^{g'}_f(x) \sum_{n=0}^\infty
  (2n+1)P_n(\mu) P_0(\mu) \phi^{g'}_n(x) d\mu'\:,
  \label{eq:fission_deriv_1}
\end{equation}
giving:
\begin{equation}
  \frac{1}{k} \sum_0^G \chi^g \nu \sigma^{g'}_f(x) \phi^{g'}_n(x)
  \delta_{n0}\:,
  \label{eq:fission_deriv_1}
\end{equation}
where $\chi^g$ and $\sigma^{g'}_f\phi^{g'}$ means that an effective
inner product amongst group terms will be computed with the discrete
$\chi$ spectrum and the fission reaction rate. In a multigoup form, if
we choose to represent the energy dependent moments as vectors,
$\mathbf{\Phi_n}$, composed of all energy groups, then this inner
product takes the form of a spatially dependent fission matrix
$\mathbf{F}$:
\begin{equation}
  \mathbf{F} = \nu
  {\tiny \begin{bmatrix}
    (\chi^0 \sigma_f^0) & (\chi^0 \sigma_f^1) & \cdots & (\chi^0
      \sigma_f^G) \\
    %%
    &&&\\
    %%
    (\chi^1 \sigma_f^0) & (\chi^1 \sigma_f^1) & \cdots & (\chi^1
      \sigma_f^G) \\
    %%
    &&&\\
    %%
      \vdots & \vdots & \ddots & \vdots \\ 
    %%
    &&&\\
    %%
    (\chi^G \sigma_f^0) & (\chi^G \sigma_f^1) & \cdots & (\chi^G
      \sigma_f^G) \\
  \end{bmatrix}}\:.
  \label{eq:fission_matrix}
\end{equation}
This gives a final multigroup $P_N$ fission source of:
\begin{equation}
  \frac{1}{k} \mathbf{F} \mathbf{\Phi_n} \delta_{n0}
  \label{eq:pn_fission_source}
\end{equation}
where the spatial dependence of the flux moments and fission matrix
has been supressed.

### 2013-05-04 doc/dissertation/document/chapters/spn_equations/spn_equations.tex
adding pn fission source derivation

the integration of the scattering emission term of energy has been

@@

\subsection{Spatial Discretization of the $SP_N$ Equations}
\label{subsec:spatial_discretization}
To this point, the formulation of the multigroup $SP_N$ equations
presented have discretized the transport equation in angle and energy
but have yet to consider the spatial component of phase space. We
don't explicitly consider it in this document but instead we will
briefly comment on possible means of discretization. Of primary
interest here is the discretization of the diffusive term in
Eq~(\ref{eq:spn_multigroup_system}) and the current term on the
boundaries in Eq~(\ref{eq:spn_multigroup_bnd}). Many popular choices
for spatial discretization are available here and include finite
difference and finite element formulations. In the Denovo package at
ORNL, Evans has implemented a finite volume scheme for these
equations resulting in:
\begin{equation}
  discretization
  \label{eq:spn_spatial_discretization}
\end{equation}
Although arbitrary grids could be handled effectively
through a finite element scheme, the rectilinear grid used in Denovo
is easily discretized through the finite volume method in a
conservative form. This conservative form is ideal for the
diffusion-like equations in the domain given by
Eq~(\ref{eq:spn_multigroup_system}) and the effective boundary current
conditions given by Eq~(\ref{eq:spn_multigroup_bnd}) as the neutron
current is balanced from cell-to-cell, continuity of the flux is
preserved across cell/material boundaries, and boundary conditions are
naturally represented through cell-face currents. Furthermore, this
spatial discretization is a natural extension of the balance
principles used to arrive at the general transport equation.

\subsection{Eigenvalue Form of the $SP_N$ Equations}
\label{subsec:eigenvalue_form}

>>>>

the integration of the scattering emission term over energy has been

@@

\subsection{Spatial Discretization of the $SP_N$ Equations}
\label{subsec:spatial_discretization}
To this point, the formulation of the multigroup $SP_N$ equations
presented have discretized the transport equation in angle and energy
but have yet to consider the spatial component of phase space. Of
primary interest here is the discretization of the diffusive term in
Eq~(\ref{eq:spn_multigroup_system}) and the current term on the
boundaries in Eq~(\ref{eq:spn_multigroup_bnd}). Many popular choices
for spatial discretization are available here and include finite
difference and finite element formulations. In the Exnihilo package at
ORNL, Evans has implemented a finite volume scheme over a Cartesian
grid for these equations resulting in:
\begin{multline}
  - \mathbb{C}_{n,i+1} \mathbb{U}_{n_i+1} - \mathbb{C}_{n,i-1}
  \mathbb{U}_{n_i-1} - \mathbb{C}_{n,j+1} \mathbb{U}_{n_j+1} -
  \mathbb{C}_{n,j-1} \mathbb{U}_{n_j-1} - \mathbb{C}_{n,k+1}
  \mathbb{U}_{n_k+1} - \mathbb{C}_{n,k-1} \mathbb{U}_{n_k-1}
  +\\ \sum_{m=1}^4 \Big[\mathbb{A}_{nm,ijk} +
    (\mathbb{C}_{n,i+1}+\mathbb{C}_{n,j+1}+\mathbb{C}_{n,k+1}+
    \mathbb{C}_{n,i-1}+\mathbb{C}_{n,j-1}+\mathbb{C}_{n,k-1})\delta_{nm}\Big]
  \mathbb{U}_{m,ijk} = \mathbb{Q}_{n,ijk}\:,
  \label{eq:spn_space}
\end{multline}
with the coefficients defined as:
\begin{equation}
  \mathbb{C}_{n,l+1} = \frac{2}{\Delta_l}\Bigg( \frac{\mathbb{D}_{n,l+1}
    \mathbb{D}_{n,l}}{\Delta_l \mathbb{D}_{n,l+1} + \Delta_{l+1} \mathbb{D}_{n,l}}\Bigg)\:,
  \label{eq:spn_coeff_plus}
\end{equation}
\begin{equation}
  \mathbb{C}_{n,l-1} = \frac{2}{\Delta_l}\Bigg( \frac{\mathbb{D}_{n,l}
    \mathbb{D}_{n,l-1}}{\Delta_{l-1} \mathbb{D}_{n,l} + \Delta_{l}
    \mathbb{D}_{n,l-1}}\Bigg)\:,
  \label{eq:spn_coeff_minus}
\end{equation}
where $l={i,j,k}$ along the cardinal directions of the Cartesian grid
cell in Figure~\ref{fig:mesh_cell}. The details of the derivation are
provided in Appendix~\ref{chap:spn_spatial_discretization}.

Although arbitrary grids could be handled effectively
through a finite element scheme, the rectilinear grid used in Denovo
is easily discretized through the finite volume method in a
conservative form. This conservative form is ideal for the
diffusion-like equations in the domain given by
Eq~(\ref{eq:spn_multigroup_system}) and the effective boundary current
conditions given by Eq~(\ref{eq:spn_multigroup_bnd}) as the neutron
current is balanced from cell-to-cell, continuity of the flux is
preserved across cell/material boundaries, and boundary conditions are
naturally represented through cell-face currents. Furthermore, this
spatial discretization is a natural extension of the balance
principles used to arrive at the general transport equation.

\subsection{Eigenvalue Form of the $SP_N$ Equations}
\label{subsec:eigenvalue_form}
For criticality problems, the space, angle, and energy discretizations
for the $SP_N$ equations can be applied to the general eigenvalue from
of the transport equation given by
Eq~(\ref{eq:eigenvalue_transport}). We can readily replace the fixed
source in Eq~(\ref{eq:multigroup_spn_equations}) with the fission
source derived in \S~\ref{sec:pn_eigenvalue}:
\begin{multline}
  -\nabla \cdot \Bigg[\frac{n}{2n+1}\mathbf{\Sigma_{n-1}}^{-1} \nabla
    \Big(\frac{n-1}{2n-1} \mathbf{\Phi_{n-2}} +
    \frac{n}{2n-1}\mathbf{\Phi_n} \Big) \\+
    \frac{n+1}{2n+1}\mathbf{\Sigma_{n+1}}^{-1} \nabla
    \Big(\frac{n+1}{2n+3}\mathbf{\Phi_n} +
    \frac{n+2}{2n+3}\mathbf{\Phi_{n+2}}\Big) \Bigg] \\+
  \mathbf{\Sigma_n} \mathbf{\Phi_n} = \frac{1}{k} \mathbf{F}
  \mathbf{\Phi_n} \delta_{n0} \ \ \ \ \ \ \ \ \ n = 0,2,4,\cdots,N\:.
  \label{eq:multigroup_spn_eigenvalue}
\end{multline}
with the fission matrix $\mathbf{F}$ given by
Eq~(\ref{eq:fission_matrix}).
